---
title: "Khan_data607_assignment_2"
author: "Mehdi Khan"
date: "September 8, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=65),tidy=TRUE)
```

```{r pressure, echo=FALSE}
##plot(pressure)
```
The following sql query was used to create a table in SQL Server with movies data:

use
data607

CREATE TABLE movies(
	movie_id int identity(1,1),
	movie_critic varchar(50),
	movie_name varchar(50),
	movie_rate int
);

INSERT INTO movies
VALUES	('John',  'Death Note',2),
		('John',  'Dunkirk',5),
		('John',  'Mother!',3),
		('John',  'Logan Lucky',4),
		('John',  'Wonder Woman',3),
		('John',  'Unlocked',2),
		('Angela',  'Death Note',1),
		('Angela',  'Dunkirk',4),
		('Angela',  'Mother!',3),
		('Angela',  'Logan Lucky',4),
		('Angela',  'Wonder Woman',5),
		('Angela',  'Unlocked',1),
		('Jared',  'Death Note',4),
		('Jared',  'Dunkirk',5),
		('Jared',  'Mother!',3),
		('Jared',  'Logan Lucky',3),
		('Jared',  'Wonder Woman',4),
		('Jared',  'Unlocked',2),
		('Steven',  'Death Note',2),
		('Steven',  'Dunkirk',5),
		('Steven',  'Mother!',3),
		('Steven',  'Logan Lucky',3),
		('Steven',  'Wonder Woman',5),
		('Steven',  'Unlocked',1),
		('Becky',  'Death Note',1),
		('Becky',  'Dunkirk',5),
		('Becky',  'Mother!',3),
		('Becky',  'Logan Lucky',4),
		('Becky',  'Wonder Woman',5),
		('Becky',  'Unlocked',2)


Using package RODBC to connect to SQL Server:
```{r}
library(RODBC)

```
Connecting to the data source (SQL Server) through ODBC connection:

```{r}
## the name of the DSN ("data607") and appropriate credentials are provided 
## to creare the connection

dataSrc <- odbcConnect("data607", uid="DataUser", pwd="Data607@Fall2017")

```

Redaing from the database:
```{r}
## See the accessible tables, tableType = "TABLE" is used to get only table objects
## otherwise it may list many different data objects

sqlTables(dataSrc, tableType = "TABLE" )


## get the table (in this case table called 'movies') and assign the data to a dataframe 
moviesDS <- sqlFetch(dataSrc,"movies", stringsAsFactors = FALSE)
head(moviesDS)

```
Subsetting the dataframe:
```{r}
## subsetting the data
subset(moviesDS, moviesDS$movie_rate >3)

```
USing SQL Query to load data (instead of subsetting):
```{r}
## It was also possible to get the required data instead of loading all data
## and subsetting from it. a sql query would do the job

requiredData <- sqlQuery(dataSrc, paste('Select * from "movies"', 'Where "movie_rate" > 3'))
head(requiredData)
```
The connection object needs to be closed once it is no longer needed:
```{r}
odbcClose(dataSrc)
```

 
 ALTERNATIVE APPROACH:
 Reading and loading data from a CSV file that was created as an output of a sql (select * from movies)
 to get all the data from the 'movies' table in a sql server databse
 
 
```{r}
## The csv file was stored in the working directory, so only the name of the
## file is enough (instead of providing the whole path)
csvMoviesDS <- read.csv("movies.csv", stringsAsFactors = FALSE)
head(csvMoviesDS)

## This dataframe (csvMoviesDS) now can be used like any other dataframe in R
```
Some data operations:
```{r}
library(dplyr)
filter(moviesDS, movie_rate == 1)
```
dplyr chaining;
Selecting the highest rated movies and number of times they received such rating:
```{r}
highestRated <- moviesDS %>% select(movie_name,movie_rate) %>% filter(movie_rate == 5) %>% count(movie_name, movie_rate)

names(highestRated)[3]='count'
highestRated
```

Raning the movies based on their ratings:
```{r}
## movie list by their total rating
most_loved_hated_movies <- aggregate(moviesDS$movie_rate, by=list(movie_name =moviesDS$movie_name), FUN = sum)


## sorting movies by rate
most_loved_hated_movies <- most_loved_hated_movies[with(most_loved_hated_movies, order(most_loved_hated_movies$x, decreasing = TRUE)),]

names(most_loved_hated_movies)[2]="overall_rating"

most_loved_hated_movies


```


